---
title: "DI_Demo"
author: "Lucia Michielin"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message = FALSE)
#install.packages("tidytext")
#install.packages("tidyverse")
#install.packages("janeaustenr")
#install.packages("magick") 

library(tidytext) # to work with unstructured data
library(tidyverse) # to do data wrangling
get_sentiments("bing")
library(janeaustenr) # to fetch the dataset
library(magick) # to display images
# these are the libraries we need if you do not have them already and you get an error remove the # from the install.packages()lines

```

## Introduction

This is a short demo to show how you can create a .gif that would record the sentiment evolution across Jane Austen books. To do so we need to follow these steps

1.  Create a Table where we are going to automatically extract each word in Jane Austen books.
2.  Calculate the sentiment values for subset of each chapters
3.  Plot a graph for each book that will show the sentiment values
4.  Collate the single graphs into a single gif

### 1. Create a word collection table

For each words we are going to collect information about

-   the book it belongs to
-   the line number where it could be found
-   the chapter where it could be found
-   the word itself

```{r Austen table}
AustenTable <- austen_books() %>% #create a new file named AustenTable that will extract info from austen_books
  group_by(book) %>% # group by every single book then
  mutate( # manipulate the data to create
    linenumber = row_number(), # a line number column that would count in which row the word was
    chapter = cumsum(str_detect(text, # the chapter number. We can do so by using regex and find lines starting with chapter followed by a space and a letter 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>% #  This line removes the grouping, so subsequent operations will be applied to the entire dataset rather than grouped subsets.
  unnest_tokens(word, text) # This tokenises the text column, splitting it into individual words and creating a new row for each word

head(AustenTable)
```

### 2. Create a word collection table

Now that we have the list of all words we extract the average sentiment of subsets of each chapter of each book. To do so as before we manipulate our dataset to do what we want. Because we want to create an uniform pattern that would simulate a tapestry we want to divide our subset to

```{r Austen Sentiment}
jane_austen_sentiment <- AustenTable %>% # Load Jane Austen's books dataset and start a chain of operations to create a Jane_austen_sentiment dataset
  inner_join(get_sentiments("bing"), relationship = 'many-to-many') %>% # Join the dataset with the Bing lexicon sentiment dictionary
  group_by(book, chapter) %>%   # Group the dataset by book and chapter
  mutate(index = rep(1:10, each = ceiling(n() / 10), length.out = n())) %>% # Create an index to split the chapters into 10 segments no matter how long the chapter it is 
  group_by(book, chapter, index) %>% # Regroup the dataset by book, chapter, and index
  count(sentiment) %>%  # Count the occurrences of each sentiment within each segment
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% # Reshape the data from long to wide format
  mutate(sentiment = positive - negative,index= as.factor(index))%>%  # Calculate sentiment score (positive - negative) for each segment
  filter(!chapter=="0") # Filter out chapters with the value "0" (if any)

head(jane_austen_sentiment)
```

### 3. Plot a graph for each book

Plot a graph for each book that will show the sentiment values tapestry across the book To do so we need to do a series of small steps

#### A. Create a directory to which the images will be written

```{r Austen directory}
dir_out <- file.path("outputs/Austen") # Define the directory path where the outputs will be saved
dir.create(dir_out, recursive = TRUE) # Create the directory if it doesn't already exist
```

#### B. List all the books that are inside our dataset

```{r Austen List of Books}
books <- unique(jane_austen_sentiment$book)
books 
```

#### C. Find which is the max number of chapters each book has

```{r Books with more chapters}
most_chapter <- max(jane_austen_sentiment$chapter, na.rm = TRUE)# Find the maximum chapter number in the dataset 'jane_austen_sentiment'
most_chapter
```

#### D. Print a graph for each book using a for loop

Now we have all that we need to create a for loop that will automatically create a graph for each book. A "for" loop is a control flow statement in programming languages that allows you to repeatedly execute a block of code a specified number of times or iterate over a sequence of values.

```{r Printing For Loop}
for (y in books) {# Iterate over each book in the 'books' vector y is the name we are going to the iteration variable it can be anything as long as you are consistent
  
  p <- # p is just a name we are giving to the plot again you can change it as long as you are consistent
    jane_austen_sentiment %>%
    filter(book == y) %>% # Filter the 'jane_austen_sentiment' dataset for the current book
    ggplot(aes(chapter,index, fill= sentiment)) +  # Create a ggplot object with chapter, index, and sentiment as aesthetics
    geom_tile() +# Add a tile layer to create a heatmap
    scale_x_continuous(breaks=seq(1,most_chapter,1), expand = c(0,0))+  # Customise x-axis scale to show breaks from 1 to 'most_chapter'
    scale_fill_gradient(low="blue", high="red", limits = c(-20, 40))+# Customise fill scale to use a gradient from blue to red
    theme_bw()+ # Apply a black-and-white theme
    guides(fill="none")+  # Remove the fill legend
    ggtitle(y)+ # Add a title to the plot with the current book's name
    coord_fixed(ratio = 1, ylim = c(10,1), xlim = c(0.5,most_chapter+0.5))+  # Fix the aspect ratio and set limits for y and x axes
    theme(  # Customize theme to remove y-axis labels and ticks
      axis.title.y = element_blank(),
      axis.text.y= element_blank(),
      axis.ticks.y = element_blank()
    )
  
  fp <- file.path(dir_out, paste0(y, ".png"))# Define the file path where the plot will be saved
  
  ggsave(plot = p,  # Save the ggplot object as a PNG file
         filename = fp,   # File path where the plot will be saved
         device = "png", # Output device type (PNG format)
         width=3500,# Width of the output in pixels
         height = 1000, # Height of the output in pixels
         units = "px") # Units of width and height (pixels)
  
}# Close the loop
```

The bit of code below it is just to look at one of the plots created

```{r Look at one of the plots}
Image <- image_read('outputs/Austen/Emma.png')
Image
```

Good! We are almost there now we need to create a gif out of the single plots

### 4. Create and Save a gif

#### A. List file names and read in

```{r Look what is inside the folder}
imgs <- list.files(dir_out, full.names = TRUE) # List all file names in the directory 'dir_out' and store them in 'imgs'
img_list <- lapply(imgs, image_read)  # Read each image file from the list of file names using 'image_read' and store them in 'img_list'
```

#### B. Join the images together

```{r join the images}
img_joined <- image_join(img_list) # Join the list of images into a single animated image using 'image_join'
```

#### C. Animate 1 frame per second

```{r Animate 1 frame per second}
img_animated <- image_animate(img_joined, fps = 1) # Create an animated image from the joined image with a frame rate of 1 frame per second using 'image_animate'
```

#### D. Save to disk

```{r Save to disk}
image_write(image = img_animated,
            path = "outputs/austen.gif")
```

Let's look at what we have done

```{r look at the gif}
img_animated
```


# The End

This is the end of this demo. 
Feel free to play around with it and modify it as you want 